---
title: 'Beyond the Boxscore: Applying Team and Individual Performance Evaluation Metrics to Canadian Basketball'
author: "David Awosoga"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

The increased usage of statistical tools and methodologies in basketball has led to
significant improvements in player and team performance evaluation. This quantitative
analysis has produced many statistical metrics that provide teams with comparative advantages
over their opponents. These metrics are readily available in many American
and International basketball leagues but are particularly underutilized at many levels of
basketball in Canada, even though similar raw data is collected through mediums such
as boxscores and play-by-play data. This paper applies some of these boxscore-based
metrics to the Canadian Elite Basketball League and University Sports Basketball, with
an accompanying website [1] that makes these statistics publicly accessible. Following
a brief history of the advancements of analytics in basketball and background on the
boxscore and its uses, the Four Factors of Basketball will be used to derive expected win
percentage models for teams over the course of a season. The introduction of these models
will be followed by discussions on their statistical background and characteristics, and
case studies will illustrate unique findings resulting from their construction. Advanced
individual player metrics such as Box Plus Minus, Player Efficiency Rating, Value over
Replacement Player, and Points per Possession will also be calculated to improve our
notion of player value within these leagues.

Keywords : Sports statistics, Four Factors, Canadian Basketball Analytics

# 1 Background

Basketball statistics began with tabulating points scored, personal fouls, and free throws and
dividing these aggregate totals by the number of games played to produce *per-game statistics*.
These statistics are easy to compute, widely understood by the general public, and still used
by sports broadcasters and fans alike. However, they hold very little value to analysts when
describing player performance because they fail to consider playing time, making them poor
indicators of efficiency and productivity. Introducing basic *percentages* to measure shooting
proficiency resolved some differences in usage rates but gave way to untrustworthy sample
sizes and did not properly account for the lower-percentage but increased value of the 3-point
shot. As new metrics were introduced to try and combat these deficiencies, ambiguity arose
about their context and employment and little headway in basketball analytics was made until
the late 1990s. It was then that *regression-based statistics* were introduced to compare the
predictive power and standard error of models and per-game statistics were succeeded by the
more stable and informative *per-minute statistics*. *Possession-based* metrics introduced a new
wave of analytics that provided a strong correlation with wins on a team level and could be
broken down for individual players as well [2]. Since teams alternate possessions in a game,
playing basketball was abstracted to an optimization problem of how to maximize a team’s
per-possession efficiency [3]. Current leaders in advanced NBA analytics involve plus-minus
statistics [4], such as Box Plus-Minus (BPM) [5], Estimated Plus-Minus (EPM), and Daily
Plus-Minus (DPM). Other metrics with more exotic acronyms have experienced increased
popularity in recent years, including LEBRON, a player role and luck-adjusted metric from
BBall Index [6], and RAPTOR, the Robust Algorithm (using) Player Tracking (and) On/Off
Ratings developed by Nate Silver of FiveThirtyEight [7].

# 1.1 Data Collection and Structure

Boxscores are the most accessible source of detailed game analysis at all levels of basketball.
They are compact tables that contain the summarized raw counting statistics of players, which
are then aggregated to comprise team performance, giving people both a cursory and in-depth
review of a game’s happenings with greater detail than just viewing the final score. However,
boxscores do not tell the whole story, and they have been succeeded by more informative data
sources such as play-by-play and player tracking. Play-by-play data refers to a game report that
gives the sequential listing of events that determine a game’s outcome, chronicling the score and
players on the court at the time of any event occurrence. They contain useful information for
predicting the outcome of a particular matchup and simulating the progression of events that
can lead to such an outcome [8]. The introduction of spatio-temporal data via player tracking
and wearable technologies in the 2010s has revolutionized basketball analytics, as cameras
are now able to determine the locations of all the players on the court, as well as the (x,y,z)
coordinates of the ball [9] with incredible accuracy and precision. This has given researchers
the ability to determine the relationship between shot selection and opponent location, as well
as provide improved analysis of individual defensive performance [10] - something historically
undervalued by boxscore and play-by-play-based statistics. Bayesian regression models and
Markov Chains have been used to model player career projections and team evaluations, leading
to the creation of new metrics [11].

The leagues under consideration are University Sports (**U SPORTS**), the governing body
and leader of university sport in Canada [12], and the Canadian Elite Basketball League
(**CEBL**), the top domestic league in Canada. U SPORTS player and team data is taken from
‘Coach’s View’ Season Summary Statistics that are publicly available for each basketball team
by PrestoSports [13]. Information about player position, which is incorporated in individual
performance evaluation metrics, is taken from usportshoops [14]. Statistics from the 2015-2016
through 2019-2020 seasons were used as the training set for created models, and the 2021-2022
season was used as the test set for predictions (the 2020-2021 season was cancelled due to
the covid-19 pandemic). The CEBL played its inaugural season in 2019 and is comprised of
talent from leagues such as the NBA G-League, USPORTS, the NCAA, and multiple high-level
international leagues. Season summary data and roster information for the CEBL is taken from
RealGM [15]. Since neither league uses player tracking and play-by-play data can be difficult to
parse efficiently, the season summaries were taken exclusively from game boxscores. The data
was collected using web-scraping in the R [16] programming language, and scripts were written
to calculate advanced statistics from these basic statistics and output them in easy-to-read
Microsoft Excel tables.


# 1.2 Team Evaluation Using the Four Factors of Basketball
The Four Factors of Basketball were introduced by statistician and basketball analytics pioneer
Dean Oliver in his 2003 book Basketball on Paper [17]. They are based on the four main events
that affect an offensive possession in basketball - a field goal attempt, a turnover, an offensive
rebound, and a foul. While other more obscure events can affect or terminate a possession,
the four factors capture the overwhelming majority of possession-ending occurrences, making
models that utilize them relatively simple and incredibly informative. Since these factors can
be expressed in terms of percentages they are comparable across seasons and even eras of
basketball with some slight modifications and estimations.

* *Effective Field Goal Percentage (eFG%)* calculates how efficiently a team scores,
adjusting for the extra value gained from a 3-point shot versus a 2-point shot. $eFG\% = FG+0.5(3P)/FGA$
* *Turnover Percentage (TOV%)* measures the percentage of a team’s possessions that
end in a turnover. Since possessions are not officially tracked in U SPORTS boxscores,
estimates with good empirical accuracy are used in their place: $POSS ≈ FGA+0.44×FTA − OFF + TO$ [3]. $TOV\% = TO /POSS$
* *Offensive Rebound Percentage (ORB%)* is defined as the percentage of offensive
rebounds that a team obtains out of total available rebounds from missed field goal
attempts and free throws. $ORB\% = OFF / TOT$
* *Free Throw Factor (FTF%)*, the percentage of “free" points a team gets per field goal
attempt. It underscores the conventional wisdom that the ability to get to the line does
not have much bearing if a team is not proficient at making free throws. $FTF\% = FT/FGA$

The rest of the paper is organized in the following manner. The Four Factors of Basketball
will be used in Section 2 on U SPORTS team data to model a team’s expected win percentage,
recalculating the coefficients to better understand each factor’s impact on the men’s and
women’s games. Section 3 introduces techniques to quantify individual player performance in
the CEBL by extending aggregate methods and reports the results, which are summarized in
Section 4.

# 2 Expected Win Percentage
The expected wins model modifies baseball’s “Expected Runs" metric [18] by regressing a
team’s wins on the four factors, a useful application of which is the ability to determine the
“importance" of each factor. Once the factors have been ranked, a strategic framework can be
devised within a team’s operations to properly allocate time to developing skills that better
align with the significance of the factor in question. By adhering to a regimented training
model based on the four factors and their relationship to winning, teams can manipulate their
factor scores to overcome talent constraints and optimize their expected win percentage.

# 2.1 Methodology
A multiple linear regression is a common empirical research tool that attempts to explain a
measurable outcome (the dependent variable, $Y_i$) using well-understood determinants (independent
variables), while accounting for random error $ϵ$. Independent variables are represented
as the set $X = {X_m,m ∈ 0, 1, 2, . . .M}$, where $X_0 ≡ 1$ to provide the traditional intercept
(constant) term. The general form of a multiple regression model is then defined as
$$Y = β_0 + β_1X_1 + β_2X_2 + . . . + β_MX_M + ϵ$$
such that
$$ E(Y |X_1, . . . ,X_M) = β_0 + β_1X_1 + β_2X_2 + . . . + β_MX_M$$
Multiple regression model assumptions [19] must hold so that the parameters $β_0, β_1, . . . , β_M$ of the model can be estimated using Ordinary Least Squares Estimators. Ordinary least squares
estimators minimize the sum of squared errors for each parameter, represented as

$$ S(β_m) = \sum_{i=1}^{N} (y_i − E(y_i))^2 = \sum_{i=1}^{N} (y_i − β_0 − β_1x_{i1} − . . . − β_Mx_{iM})^2$$
With these assumptions satisfied, this process will result in the estimators $β_m$ being the
best linear unbiased estimators of the parameters. If the error term of the model is normally
distributed, then the least-squares estimators and the dependent variable $y_i$ for each observation
will also be normally distributed. The coefficient of determination, more commonly referred
to as a model’s R^2^ value, will be used in the following discussions to demonstrate how well the
four factors predict team success in a season. Adjusted R^2^ ($\overline{R}^2$) is an alternative measure of
goodness-of-fit that accounts for some deficiencies present within the basic R^2^ construct and
penalizes the inclusion of variables that do not add value to a model’s predictive ability.

# 2.2 Model Specification
Because U SPORTS teams in different conferences play a different number of regular season
games, expected winning percentage is used as the response variable in place of expected
wins so that the model is properly generalized across the league. To use the four factors as
explanatory variables to model a team’s expected win percentage, their offensive and defensive
factor performances are split into two categories. The offensive factors are denoted by their
familiar name, while the factors of their opposition are prepended with “opp", for eight total
variables. Since the actual win percentage is a value ranging from 0 to 100 inclusive and is
linear in nature, there is no guard against expected values outside the realm of possibility, such
as values over 100% or negative percentages. Due to its propreitary nature it is unclear whether
Oliver’s model included an intercept term or not, and the terms’ effects are worth considering.
The inclusion of a constant term often has an inconsequential interpretation, but excluding it
could have significant ramifications on a model’s R^2^ and normality assumptions. Therefore,
models with and without the constant term will be generated in the following manner and
compared during analysis:

$$ WinPecentage = β_0 + β_1eFG\% + β_2TOV\% + β_3ORB\% + β_4FTF\%
+ β_5oppeFG\% + β_6oppTOV\% + β_7oppORB\% + β_8oppFTF\%$$

$$ WinPecentage =  β_1eFG\% + β_2TOV\% + β_3ORB\% + β_4FTF\%
+ β_5oppeFG\% + β_6oppTOV\% + β_7oppORB\% + β_8oppFTF\%$$

# 2.3 Analysis
Table 1 report the output of the regressions run from equations 1 (*M<sub>1<sub>* and *W<sub>1<sub>*) and 2 (*M<sub>2<sub>* and
*W<sub>2<sub>*), courtesy of the stargazer package from R [20]. The signs on each coefficient align with
a basic understanding of the game of basketball: shooting, rebounding, and free throws will
have a positive impact on a team’s success, while turnovers will have a negative effect. The
opposite intuitively holds for the opposition and is reflected by the coefficients on the “opp"
variables. Each coefficient is statistically significant at the 1% level except for oppFTF in *M<sub>2<sub>*,
which has a p-value of 0.012.

```{r, echo= TRUE}
# Install and Load Packages 
if (!require("pacman")) install.packages("pacman")
# Use pacman to load add-on packages as desired
pacman::p_load(pacman, caret, lars, tidyverse, rio, scales, tibble, dplyr, stringr, rvest, tidyr, knitr, ggplot2, olsrr)
oldw <- getOption("warn")
options(warn = -1)

# Set Working Directory to where files are stored !!

## Importing Data with Rio
USPORTS_Men <- import("USPORTS Five Year Four Factors Training Set M.csv")
USPORTS_Women <- import("USPORTS Five Year Four Factors Training Set W.csv")

## Create linear models with and without coefficients to compare fits

M1 <- lm(WinPercentage ~ eFG + TOV + ORB + FTF + oppeFG + oppTOV + oppORB + oppFTF, data = USPORTS_Men)
M2  <- lm(WinPercentage ~ eFG + TOV + ORB + FTF + oppeFG + oppTOV + oppORB + oppFTF - 1, data = USPORTS_Men)
W1 <- lm(WinPercentage ~ eFG + TOV + ORB + FTF + oppeFG + oppTOV + oppORB + oppFTF, data = USPORTS_Women)
W2  <- lm(WinPercentage ~ eFG + TOV + ORB + FTF + oppeFG + oppTOV + oppORB + oppFTF - 1, data = USPORTS_Women)

summary(M1)
summary(M2)
summary(W1)
summary(W2)

```

In *M<sub>1<sub>*, the adjusted R^2^ of 0.868 means that the explanatory variables account for nearly 87%
of the variation in the response variable (win percentage) of the model. This is extremely high
and comparable with results from NBA datasets [21][22], providing strong evidence in defence
of the factors’ relationship to winning. The minuscule difference between the original and
adjusted goodness-of-fit values demonstrates how little penalty each added variable receives.
Since the factors and dependent variable can be expressed with percentages, they have a very
straightforward interpretation. For example, a 1 percent change in eFG% is expected to change
a team’s win percentage by 2.333 percent, and a team that reduces its TOV% by 1 percent is
expected to enjoy a 1.962 percent increase in win percentage. *M<sub>2<sub>* omits the constant term and
consequently produces some interesting results. To start, the coefficient of determination in *M<sub>2<sub>* experiences a substantial increase to 0.974, nearly 11 points greater than that of *M<sub>1<sub>*. Another
observation from comparing the two models is the defensive emphasis of *M<sub>1<sub>* compared to the
more offense-heavy focus given to the coefficients of *M<sub>2<sub>*. The values of eFG% and oppeFG%
are seemingly flipped between the models, and rebounding also experiences a similar switch
in coefficients on both sides of the ball, from 0.695 to 0.936 offensively and −0.97 to −0.596
on defense. Some coefficients remain relatively unchanged, such as oppTOV (2.062 vs 2.022)
and oppFTF (−0.544 vs −0.479). The intercept term β0 is incorporated to ensure that the
mean of the residuals will be zero and protect the model against bias, part of the model assumptions
referenced in Section 2.1. Even though the histograms of the residuals of *M<sub>1<sub>* and *M<sub>2<sub>* have slightly different shapes as shown in Figure 2.3, those of *M<sub>2<sub>* are still very normal in their distribution, with a mean of 0.000555. Normality tests such as that of Shapiro-Wilk,
Kolmogorov-Smirnov, and Anderson-Darling all report p-values greater than 0.05 (0.084, 0.85,
and 0.33, respectively), so any uncertainties about *M<sub>2<sub>*’s legitimacy can be quickly put to rest.

```{r, echo=TRUE}

hist(residuals(M1))  # Histogram of residuals
hist(residuals(M2))  # Histogram of residuals
print(paste0("Mean of Residuals: ",round(mean(resid(M2)), 6)))
ols_test_normality(M2)

```

The women’s regression output *W<sub>1<sub>* was similar to the men’s model, with intuitively correct coefficients and an excellent R^2^ value of 0.904. The F-statistic was also very high and slightly outperformed the men’s model, giving reason to believe that the four factors may better model the women’s game. *W<sub>2<sub>* exhibits a comparable offensive emphasis to *M<sub>2<sub>* when estimating the factor coefficients and experiences a large increase in R^2^ with almost zero penalization from the dependent variables. The ability to explain nearly 98% of the model’s variation is substantial and further proves that the four factors are transferable across multiple levels of basketball. It is interesting to note that even though the average eFG% in U SPORTS basketball is vastly different between men and women (48% to 41.3%), the coefficient on these terms for both models is similar.

## 2.4 Predictive Ability

The models used the games played during the 2021-22 season to test their predictive abilities
and were compared in Table 2 using their R^2^, root-mean-square error (RMSE), and residual
square error (RSE). Even though the models without intercepts had superior R^2^ values to those
with them in the regression output with the training set, they were very similar in predicting
win percentage on the test data set. The men’s and women’s models produced nearly identical
results in explaining the variation of win percentage in the test set, but the RSEs of the women’s
models were lower than *M<sub>1<sub>* and *M<sub>2<sub>*, with *W<sub>1<sub>* being the only model to dip under 17%. Figure
1 compares the expected win percentage from *M<sub>1<sub>* (horizontal axis) and actual win percentage
(vertical axis) for the 48 men’s basketball teams during the 2021-2022 season. Within this
figure lie various applications of the four factors, two of which are explored in the following
case studies.

```{r, echo=FALSE}

# Set Working Directory to where files are stored !!

# Make Predictions!
set.seed(123)
train_women.data <- USPORTS_Women
test_women.data <- import("USPORTS Four Factors Testing Set W.csv")
train_men.data <- USPORTS_Men
test_men.data <- import("USPORTS Four Factors Testing Set M.csv")
Teams <- as.data.frame(read.csv("USPORTS Teams List.csv"))

predictions_women1 <- W1 %>% predict(test_women.data)
predictions_women2 <- W2 %>% predict(test_women.data)
predictions_men1 <- M1 %>% predict(test_men.data)
predictions_men2 <- M2 %>% predict(test_men.data)

## Compare Model Performance on 2021-22 Season Using, R^2, RMSE, and Residual Standard Error (RSE)

errors <- round(as.data.frame(matrix(c(
R2(predictions_men1, test_men.data $ WinPercentage),
RMSE(predictions_men1, test_men.data $ WinPercentage),
sigma(M1)/mean(test_men.data$WinPercentage),

R2(predictions_men2, test_men.data $ WinPercentage),
RMSE(predictions_men2, test_men.data $ WinPercentage),
sigma(M2)/mean(test_men.data$WinPercentage),

R2(predictions_women1, test_women.data $ WinPercentage),
RMSE(predictions_women1, test_women.data $ WinPercentage),
sigma(W1)/mean(test_women.data$WinPercentage),

R2(predictions_women2, test_women.data $ WinPercentage),
RMSE(predictions_women2, test_women.data $ WinPercentage),
sigma(W2)/mean(test_women.data$WinPercentage)
), ncol = 4)), 3)

colnames(errors) <- c("M1", "M2", "W1", "W2")
rownames(errors) <- c("R^2", "RMSE", "RSE")
kable(errors, caption = "Comparing Predictive Abilities of Each Model")

```

## 2.5 Case Study: Measuring Statistical Accomplishment

One application of the expected win percentage model is using the predicted values as measures
of statistical dominance and inferiority. Here, the “best" and “worst" statistical teams
are defined as those with the highest and lowest expected win percentages, the extreme values
from the x-axis shown in Figure 1. Coming as little surprise to anyone familiar with U
SPORTS men’s basketball over the past decade, the best statistical team during the 2021-
2022 season were the eventual national champion Carleton Ravens. They were the cream of
the crop amongst a deep pool of talented teams, with an expected win percentage nearly 20
points greater than the next best school, cruising to a 14-0 regular season. Statistically, Carleton
boasted the 2nd best eFG%, 2nd best ORB%, and the 2nd best opponent ORB% in U
SPORTS, at 53.7%, 38.2%, and 20.9%, respectively. However, their defensive performance is
what warranted their infeasible expected winning percentage of 111%. Their 37.8% opponent
eFG was the lowest recorded in U SPORTS men’s basketball, not only in the season of interest
but across the entire 6-year data set. Only four other teams, including the 2017-2018 Ravens,
had recorded a season allowing an opponent eFG under 40% since the 2015-2016 season. When
examining statistical inferiority among men’s U SPORTS teams, one does not have to look any
further than the Grant Macewan University Griffins. The Griffins failed to win a game during
the regular season and their prohibitively poor performance resulted in them achieving an expected
win percentage of -8%, the only predicted value below zero among men’s results. Shown
in Table 3, their 40.4% eFG% was the lowest in U SPORTS in the 2021-2022 season, and the
second-worst output across the entire 6-year data set. Defensively, Macewan’s opponent eFG%
of 56.9% was the highest surrendered in the six years under consideration, nearly 1% worse
than the next closest team, the winless 2017-2018 Trinity Western Spartans. The Griffins
were among the bottom-five teams in nearly every other statistical category during 2021-2022,
including the 3rd-worst TOV%, 4th-worst ORB% and FTF%, and 2nd-worst opponent ORB%.
An interesting thing to note is that Macewan had the 3rd-best opponent FTF% in the country,
and this detail demonstrates why FTF% is the least valuable of the four factors - excellence in
this category cannot overcome poor results among the other factors.

# 2.6 Case Study: Assessing Achievement Using Residuals

Another byproduct of the predicted model construction in Figure 1 is that the difference between
expected and actual win percentage gives the residuals, visually described as the vertical
distance between the $45^◦$ line and the point of interest. Taking the maximum and minimum
values of these residuals yields a unique interpretation - teams with the greatest under/overachievement.
Across the men’s teams, Acadia University had the dubious distinction of U
SPORTS’ most underachieving team. The lowly Axemen finished dead last in their conference
with a 6.25 win percentage (1 win, 15 losses) but had an expected win percentage nearly 18
points higher, at 24% (4 wins and 12 losses). Acadia had solid offensive production with a
49.6% eFG% that was good for 16th in the country, and coupled this shooting proficiency with
an above-average 18.1% TOV%, 20th in U SPORTS. This set the precedence for perceived
success but is where the positive reviews come to a crashing halt. The Axemen were ranked
41st in the country in opponent eFG%, 44th in opponent FTF%, and 46th in ORB% at a paltry
21.4%. However, there is more than meets the eye about Acadia’s undisciplined defense, a
detail that provides clarity to Acadia’s win percentage discrepancy. The Axemen received 13
disqualifications during the regular season, the third highest total in U SPORTS. This statistic
is not shown by the four factors and occurs when a player is ejected from a game for committing
five personal fouls. The early exits of players who were key offensive contributors cost Acadia
on multiple occasions late in games, with five of their losses coming by 6 points or less. On the
opposite end of the achievement spectrum, the McGill Redbirds were an expected 8-win 4-loss
team that took advantage of a weak conference schedule en route to a perfect 12-0 record,
winning the RSEQ division. The Redbirds’ 50% difference between their win percentage and
that of the second place team (Concordia, 6-6) was the largest such gap in any division in
U SPORTS. Additionally, their division had the smallest divisional parity index as per win
percentage standard deviation ratios [23], an indicator of the overall closeness of talent within
that division. McGill achieved success with an elite defensive scheme, ranking 5th overall in
opponent FTF%, 6th in opponent TOV%, and 11th in opponent eFG%. This staunch defense
won the Redbirds multiple "close" regular season games - 9 of their 12 victories came by 6
points or less. However, their modest 49.5% eFG% and below-average 21.9% TOV% proved to
be insurmountable against the stronger out-of-conference opponents that McGill faced in the
USPORTS Final 8. In the quarterfinal, they were overwhelmed by the statistically superior
and eventual U SPORTS bronze medallists Alberta Golden Bears and then fell to the Canada
West Champion Victoria Vikes in the consolation semifinal. This marked a disappointing end
to McGill’s once hopeful season, summarized by a lack of offensive prowess that was hidden
behind the relative weaknesses of their regular season opponents.

```{r, echo = FALSE}
pacman::p_load(pacman, caret, lars, tidyverse, rio, scales)
data_mod_men <- data.frame(Predicted = predictions_men1, Observed = test_men.data$WinPercentage, Name = Teams$Name)  # Create data for ggplot2

ggplot(data_mod_men,                                     # Draw plot using ggplot2 package
       aes(x = Predicted, y = Observed, label = Name)) +
  geom_point(color = ifelse(data_mod_men$Name == "Acadia" | data_mod_men$Name == "Carleton" | data_mod_men$Name == "MacEwan" | data_mod_men$Name == "McGill", "black" ,"#f8766d"))+
  geom_abline(intercept = 0,
              slope = 1,
              color = "#00bfc4",
              size = 1) + 
  geom_text(aes(
                label=ifelse(
                  Name == "McGill" | Name == "Acadia" | Name == "Carleton" | Name == "MacEwan",
                  as.character(Name),'')),hjust=0.5, vjust=-0.6) + xlab(
                    "Expected Win Percentage") + 
  ylab("Actual Win Percentage") + 
  scale_x_continuous(labels = percent, breaks = seq(-0.25,1.20,0.25)) + 
  scale_y_continuous(labels = percent) +
  ggtitle("Men") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5))

```

# 3 Individual Player Evalution
A question that arises following the effective performance of advanced analytical tools in evaluating
team performance is whether these methods can be extended to serve as proxies for
quantifying an individual player’s talent. Such proxies provide great value to coaching staff
for scouting reports and playing time allocation, and to fans and other media members in
making informed year end league award nominations. Individual player metrics generally fall
into two categories, the first being statistics that measure a player’s value in terms of their
contributions to their team’s success in a specific area, a top-down approach. This category
takes possession-related statistics such as the four factors and points per possession (PPP) and
disaggregates them so that player statlines can be used to calculate individual PPP, eFG%,
TOV%, and FTF%. This technique used in paid applications such as Synergy Sports Technology
[24] and provides insight into identifying the relative strengths (and weaknesses) of
each player to properly contextualize their valuation. Additionally, play-by-play and spatial
data make it possible to determine a player’s ORB%, which cannot be computed using only
boxscore numbers. The other branch of individual player metrics are called “catch-all" statistics,
a bottom-up approach that aims to summarize observable contributions of a player and
conglomerate them into one number to determine their value. Catch-all metrics are extremely
useful for overall comparisons of players but are a hotly contested subject area in NBA circles
due to the relative secrecy of their formulation. However, some of these calculations have been
publicized, and three of them; Player Efficiency Rating (PER) [25], Box-Plus Minus (BPM)
[5], and Value over Replacement Player (VORP) [5], will be discussed in greater depth.

# 3.1 Catch-All Statistics
PER was created by former ESPN columnist John Hollinger and was the first all-in-one statistic
to attempt to quantify a player’s value. It takes a players’ boxscore contributions: field goals,
free throws, 3’s, assists, rebounds, blocks and steals, turnovers, missed shots, and fouls, adding
the positive contributions and penalizing the negative ones using detailed formulas [26]. It
adjusts for per-minute productivity and team pace, as well as the league averages in all the
statistical categories, and is normalized such that a PER of 15 is considered league average. It
is biased towards offensive output since these totals are better recorded in boxscsores so players
with an unseen defensive impact will be considered underrated overall, but as the first of its kind
PER revolutionized player evaluations during the 2010s. BPM is the successor to PER and has
replaced it in most NBA front offices. Created by Daniel Myers, it is much more comprehensive
and better-rounded metric than its predecessor (though still a bit offensively biased). BPM
estimates a player’s contribution in points above league average per 100 possessions played. It
is produced using similar boxscore-based statistics as PER, but also includes a player’s position,
offensive role, and their teams’ overall performance in the even more comprehensive calculations
[27]. BPM is set such that league average is 0.0 (0 points above or below average). For example,
a BPM of 5.0 means that a player’s team is 5 points per 100 possessions better with them on
the floor than with average production from another player. VORP converts BPM, which
does not take playing time into account, into an estimate of a player’s overall contribution to
their team. It is measured against what a theoretical bench/replacement player would provide
(someone with a BPM of -2.0 by definition), by including the number of possessions that a
player has featured in and the fraction of their teams’ games that they have played. VORP
removes any biases in pace and is considered by many as one of the better measures of actual
value contributed to a team [28].


```{r echo = FALSE, results='asis'}
#The numbers here will be slightly different than the website because the season is still in progress

yr <- 2022

# Set Global Variables

Positions <- data.frame(2.130, 8.668, -2.486, 0.992, -3.536, 1.667, 50)
colnames(Positions) <- c("Intercept","% of TRB", "% of STL", "% of PF", "% of AST", "% of BLK", "Min Wt")
rownames(Positions) <- "Modern"
OffensiveRole <- data.frame(6, -6.642, -8.544, -0.330)
colnames(OffensiveRole) <- c("Intercept", "% of ast", "% of Threshold", "Pt Threshold")
DefaultPos <- 4

BoxPlusMinusCoefficients <- data.frame(0.860,	-0.560,	-0.246,	0.389,	0.580,	-0.964,	0.613,	0.116,	0.000,	1.369,	1.327,	-0.367)
BoxPlusMinusCoefficients <- rbind(BoxPlusMinusCoefficients, c(0.860,	-0.780,	-0.343,	0.389,	1.034,	-0.964,	0.181,	0.181,	0.000,	1.008,	0.703,	-0.367))
colnames(BoxPlusMinusCoefficients) <- c("Adj. Pt",	"FGA",	"FTA",	"3 Pt FG (bonus)",	"AST",	"TO",	"ORB",	"DRB",	"TRB",	"STL",	"BLK",	"PF")
rownames(BoxPlusMinusCoefficients) <- c("Pos 1", "Pos 5")

OffensiveBoxPlusMinusCoefficients <- data.frame(0.605,	-0.330,	-0.145,	0.477,	0.476,	-0.579,	0.606,	-0.112,	0.000,	0.177,	0.725,	-0.439)
OffensiveBoxPlusMinusCoefficients <- rbind(OffensiveBoxPlusMinusCoefficients, c(0.605,	-0.472,	-0.208,	0.477,	0.476,	-0.882,	0.422,	0.103,	0.000,	0.294,	0.097,	-0.439))
colnames(OffensiveBoxPlusMinusCoefficients) <- c("Adj. Pt",	"FGA",	"FTA",	"3 Pt FG (bonus)",	"AST",	"TO",	"ORB",	"DRB",	"TRB",	"STL",	"BLK",	"PF")
rownames(OffensiveBoxPlusMinusCoefficients) <- c("Pos 1", "Pos 5")

Teams <- c("Niagara River Lions",
"Fraser Valley Bandits",
"Hamilton Honey Badgers",
"Saskatchewan Rattlers",
"Guelph Nighthawks",
"Edmonton Stingers",
"Montreal Alliance",
"Newfoundland Growlers",
"Scarborough Shooting Stars",
"Ottawa Black Jacks")

IDs <- c(1731, 2051, 2049, 2050, 2052, 2053, 1280, 2511, 2510, 2187)
team = matrix(data = c(Teams,IDs), nrow = 10)
if (yr < 2022) {
  team <- team[-which(team[,1] == "Montreal Alliance"),]
  team <- team[-which(team[,1] == "Newfoundland Growlers"),]
  team <- team[-which(team[,1] == "Scarborough Shooting Stars"),]
} 
if (yr < 2020) {
  team <- team[-which(team[,1] == "Ottawa Black Jacks"),]
}
efficiency <- paste0("https://basketball.realgm.com/international/league/128/Canadian-Elite-Basketball-League/team-stats/", yr, "/Advanced_Stats/Team_Totals/team_name/desc")

lgPace <- 0
lgThree <- 0
lgFT <- 0
lgPF   <- 0
lgFTA  <- 0
lgAST  <- 0
lgFG   <- 0
lgPTS  <- 0
lgFGA  <- 0
lgORB  <- 0
lgTO   <- 0
lgTRB  <- 0
lgMin  <- 0
lgBLK  <- 0
lgSTL  <- 0
lgPoss <- 0

lgRankings <- data_frame()
league <- data.frame()
totals <- data.frame()
TeamStats <- data.frame()

#MASSIVE FOR LOOP

for(i in 1:length(team[,1])) {
#Get url page formatted

pos <- str_locate_all(pattern = " ", team[i,1])
urlname <- ""
starting <- (length(pos[[1]])/2) %% 2
for(idx in 1:(((length(pos[[1]])/2))+1)) {
  if(idx == (length(pos[[1]])/2)+1) {
    segment <- substr(team[i,1],starting+1,nchar(team[i,1]))
  } else {
    segment<- paste0(substr(team[i,1],starting-(length(pos[[1]])/2) %% 2,pos[[1]][idx]-1), "-")
}
urlname <- paste0(urlname, segment)
starting <- pos[[1]][idx]
}
for (x in 1:length(urlname)) {
  urlname[x] <-  str_replace_all(urlname[x], "[\\s]", "")
}


gameUrl <- paste0("https://basketball.realgm.com/international/league/128/Canadian-Elite-Basketball-League/team/", team[i,2], "/",urlname, "/stats/", yr,"/Totals/All/All/player/All/desc")
roster <- paste0("https://basketball.realgm.com/international/league/128/Canadian-Elite-Basketball-League/team/", team[i,2], "/",urlname, "/rosters/",yr)
webpage1 <- read_html(gameUrl)
webpage2 <- read_html(roster)
webpage3 <- read_html(efficiency)
shotchart <- "td"
shots_html <- html_nodes(webpage1, shotchart)
roster_html <- html_nodes(webpage2, shotchart)
efficiency_html <- html_nodes(webpage3, shotchart)
shots_data <- html_text(shots_html)
roster_data <- html_text(roster_html)
efficiency_data <- html_text(efficiency_html)

## get rid of some fluff
for (x in 1:length(shots_data)) {
  shots_data[x] <-  str_replace_all(shots_data[x], "[\n\t]", "")
  shots_data[x] <-  str_replace_all(shots_data[x], ",", "")
}

for (x in 1:length(roster_data)) {
  roster_data[x] <-  str_replace_all(roster_data[x], "[\n\t]", "")
}


#Create Totals Table
idx <- which(shots_data == "Team Totals")
playerTotals <- shots_data[1: idx-1]
idx <- which(roster_data == "Starters")
rosterTotals <- roster_data[1:idx-1]

playerSeasonTotals <- data.frame(matrix(unlist(playerTotals), 
                                        nrow = (length(playerTotals)/23), 
                                        byrow = TRUE), stringsAsFactors = FALSE)
playerSeasonTotals <- playerSeasonTotals %>% mutate_at(c(4:23), as.numeric)
colnames(playerSeasonTotals) <- c("Rank", "Name", "Team", "GP", "MIN", 
                                  "PTS", "FGM", "FGA","FG%", "3PM", "3PA", 
                                  "3P%", "FTM", "FTA", "FT%", "OFF", "DEF", 
                                  "TRB", "AST", "STL", "BLK", "TOV", "PF")

playerSeasonTotals <- add_column(playerSeasonTotals, "2PM" = playerSeasonTotals$FGM - playerSeasonTotals$`3PM`, .before = 'FTM')
playerSeasonTotals <- add_column(playerSeasonTotals, "2PA" = playerSeasonTotals$FGA - playerSeasonTotals$`3PA`, .before = 'FTM')
playerSeasonTotals <- add_column(playerSeasonTotals, "2P%" = playerSeasonTotals$`2PM`/playerSeasonTotals$`2PA`, .before = 'FTM')
playerSeasonTotals <- add_column(playerSeasonTotals, "eFG%" = (playerSeasonTotals$FGM + (0.5*playerSeasonTotals$`3PM`))/playerSeasonTotals$FGA, .before = 'FTM')
playerSeasonRoster <- data.frame(matrix(unlist(rosterTotals), nrow = (length(rosterTotals)/9), byrow = TRUE), stringsAsFactors = FALSE)
teamEfficiency <- data.frame(matrix(unlist(efficiency_data), nrow = (length(efficiency_data)/19), byrow = TRUE), stringsAsFactors = FALSE)
teamEfficiency <- teamEfficiency %>%mutate_at(c(3:19), as.numeric)
#colnames(teamEfficiency) <- c("Rank", "Team", "TS%", "eFG%", )

#Get Roster Stats

LeagueAverageORtg <- sum(teamEfficiency$X15, na.rm = TRUE)/length(teamEfficiency$X1)
OffRating <- teamEfficiency$X15[which(teamEfficiency$X2 == team[1])]
DefRating <- teamEfficiency$X16[which(teamEfficiency$X2 == team[1])]
NetRatingAdj <- OffRating - DefRating
Pace <- teamEfficiency$X19[which(teamEfficiency$X2 == team[1])]
OffRatingAdj <- OffRating - LeagueAverageORtg
DefRatingAdj <- DefRating - LeagueAverageORtg
AverageLead <- NetRatingAdj * Pace / 100 / 2
LeadBonus <- 0.35/2*AverageLead
AdjTeamRating <- NetRatingAdj + LeadBonus
AdjOffRating <- OffRatingAdj + LeadBonus/2

#Conglomerate Team Stats
TeamPoints <- sum(playerSeasonTotals$PTS, na.rm = TRUE)
TeamFGA <- sum(playerSeasonTotals$FGA, na.rm = TRUE)
TeamFTA <- sum(playerSeasonTotals$FTA, na.rm = TRUE)
S6 <- ((BoxPlusMinusCoefficients[1,]$FTA/BoxPlusMinusCoefficients[1,]$FGA))
TmPtsPerTSA <- TeamPoints/(TeamFGA+TeamFTA*S6)
BaselinePtsPerTSA <- 1
TotalMinutes <- sum(playerSeasonTotals$MIN, na.rm = TRUE)
TeamTRB <- sum(playerSeasonTotals$TRB, na.rm = TRUE)
TeamSTL <- sum(playerSeasonTotals$STL, na.rm = TRUE)
TeamPF <- sum(playerSeasonTotals$PF, na.rm = TRUE)
TeamAST <- sum(playerSeasonTotals$AST, na.rm = TRUE)
TeamBLK <- sum(playerSeasonTotals$BLK, na.rm = TRUE)
TeamGames <- as.numeric(shots_data[which(shots_data == "Team Totals")+1])

#Gather Data for PER
idx <- which(shots_data == "Team Totals")
last <- idx+20
Statline <- (as.numeric(shots_data[idx :last]))
Statline[1] <- playerSeasonTotals$Team[1]
POSS <- (as.numeric(Statline[20])  + TeamFGA +0.44*TeamFTA-sum(playerSeasonTotals$OFF, na.rm = TRUE))

idx1 <- which(shots_data == "Opponent Totals")
last <- idx1+20
oppStatline <- (as.numeric(shots_data[idx1: last]))
oppPOSS <- (oppStatline[20]+oppStatline[6]+0.44*oppStatline[12]-oppStatline[14])

lgPoss <- lgPoss+POSS
teamPace <- (5*((POSS + oppPOSS)/2) / sum(playerSeasonTotals$MIN, na.rm = TRUE))

lgThree <- lgThree + sum(playerSeasonTotals$`3PM`, na.rm = TRUE)
lgFT <- lgFT + sum(playerSeasonTotals$FTM, na.rm = TRUE)
lgPF   <- lgPF   + sum(playerSeasonTotals$PF, na.rm = TRUE)
lgFTA  <- lgFTA  + TeamFTA  
lgAST  <- lgAST  + TeamAST 
lgFG   <- lgFG   + sum(playerSeasonTotals$FGM, na.rm = TRUE)
lgPTS  <- lgPTS  + TeamPoints  
lgFGA  <- lgFGA  + TeamFGA  
lgORB  <- lgORB  + sum(playerSeasonTotals$OFF, na.rm = TRUE)  
lgTO   <- lgTO   + as.numeric(Statline[20]) 
lgTRB  <- lgTRB  + TeamTRB
lgMin  <- lgMin  + sum(playerSeasonTotals$MIN, na.rm = TRUE)
lgBLK  <- lgBLK  + TeamBLK  
lgSTL  <- lgSTL  + TeamSTL

Statline <- append(Statline, c(POSS, oppPOSS, teamPace))
TeamStats <- rbind(TeamStats, Statline)

#Adjust for Team Shooting Context

playerSeasonTotals <- add_column(playerSeasonTotals, "TSA" = playerSeasonTotals$FGA + S6 * playerSeasonTotals$FTA, .after = 'PF')
playerSeasonTotals <- add_column(playerSeasonTotals, "Pt/TSA" = playerSeasonTotals$PTS/playerSeasonTotals$TSA, .after = 'TSA')
playerSeasonTotals <- add_column(playerSeasonTotals, "Adj. Pts" = ((playerSeasonTotals$`Pt/TSA` - TmPtsPerTSA)+BaselinePtsPerTSA) * playerSeasonTotals$TSA, .after = 'Pt/TSA')
playerSeasonTotals <- add_column(playerSeasonTotals, "Possessions" = playerSeasonTotals$MIN * Pace/40, .after = 'Adj. Pts')


playerSeasonTotals <- add_column(playerSeasonTotals, "ThreshPts" = (playerSeasonTotals$TSA * (playerSeasonTotals$`Pt/TSA`-(TmPtsPerTSA + OffensiveRole$`Pt Threshold`))), .after = 'Possessions')
TeamThreshPts <- sum(playerSeasonTotals$ThreshPts, na.rm = TRUE)

# Calculated Stats per 100 Possessions

playerSeasonTotals <- add_column(playerSeasonTotals, "Adj Pt" = playerSeasonTotals$`Adj. Pts` /playerSeasonTotals$Possessions * 100, .after = 'ThreshPts')
playerSeasonTotals <- add_column(playerSeasonTotals, "FGA_p" = playerSeasonTotals$FGA /playerSeasonTotals$Possessions * 100, .after = 'Adj Pt')
playerSeasonTotals <- add_column(playerSeasonTotals, "FTA_p" = playerSeasonTotals$FTA /playerSeasonTotals$Possessions * 100, .after = 'FGA_p')
playerSeasonTotals <- add_column(playerSeasonTotals, "3PM_p" = playerSeasonTotals$`3PM` /playerSeasonTotals$Possessions * 100, .after = 'FTA_p')
playerSeasonTotals <- add_column(playerSeasonTotals, "AST_p" = playerSeasonTotals$AST /playerSeasonTotals$Possessions * 100, .after = '3PM_p')
playerSeasonTotals <- add_column(playerSeasonTotals, "TOV_p" = playerSeasonTotals$TOV /playerSeasonTotals$Possessions * 100, .after = 'AST_p')
playerSeasonTotals <- add_column(playerSeasonTotals, "ORB_p" = playerSeasonTotals$OFF /playerSeasonTotals$Possessions * 100, .after = 'TOV_p')
playerSeasonTotals <- add_column(playerSeasonTotals, "DRB_p" = playerSeasonTotals$DEF /playerSeasonTotals$Possessions * 100, .after = 'ORB_p')
playerSeasonTotals <- add_column(playerSeasonTotals, "TRB_p" = playerSeasonTotals$TRB /playerSeasonTotals$Possessions * 100, .after = 'DRB_p')
playerSeasonTotals <- add_column(playerSeasonTotals, "STL_p" = playerSeasonTotals$STL /playerSeasonTotals$Possessions * 100, .after = 'TRB_p')
playerSeasonTotals <- add_column(playerSeasonTotals, "BLK_p" = playerSeasonTotals$BLK /playerSeasonTotals$Possessions * 100, .after = 'STL_p')
playerSeasonTotals <- add_column(playerSeasonTotals, "PF_p" = playerSeasonTotals$PF /playerSeasonTotals$Possessions * 100, .after = 'BLK_p')


#% Min

playerSeasonTotals <- add_column(playerSeasonTotals, "% of Min" = playerSeasonTotals$MIN /(TotalMinutes/5), .after = 'PF_p')

# % of Stats

playerSeasonTotals <- add_column(playerSeasonTotals, "% of TRB" = playerSeasonTotals$TRB / TeamTRB / playerSeasonTotals$`% of Min`, .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "% of STL" = playerSeasonTotals$STL / TeamSTL / playerSeasonTotals$`% of Min`, .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "% of PF" = playerSeasonTotals$PF / TeamPF / playerSeasonTotals$`% of Min`, .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "% of AST" = playerSeasonTotals$AST / TeamAST / playerSeasonTotals$`% of Min`, .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "% of BLK" = playerSeasonTotals$BLK / TeamBLK / playerSeasonTotals$`% of Min`, .before = '% of Min')

playerSeasonTotals <- add_column(playerSeasonTotals, "% of ThreshPts" = playerSeasonTotals$ThreshPts / TeamThreshPts / playerSeasonTotals$`% of Min`, .before = '% of Min')

#Estimate Positions
position <- vector()
for(name in playerSeasonTotals$Name) {
  if (name == "Walt Lemon Jr.") {
    name <- "Walt Lemon, Jr."
  }
  if (name == "Derek Cooke Jr.") {
    name <- "Derek Cooke, Jr."
  }
  psn <- playerSeasonRoster$X3[which(name == playerSeasonRoster$X2)]
  if (psn == "-") {
    psn <- "SF"
  }
  convertedNum <- switch(psn, "PG" = 1, "SG" = 2, "SF" = 3, "PF" = 4, "C" = 5, "GF" = 2.5, "FG" = 2.5, "G" = 1.5, "F" = 3.5, "FC" = 4.5)
  position <- append(position,convertedNum)
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Pos Num" = position, .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Est Pos 1" = Positions$Intercept + 
                                   (playerSeasonTotals$`% of TRB` * Positions$`% of TRB`) + 
                                   (playerSeasonTotals$`% of STL` * Positions$`% of STL`) +
                                   (playerSeasonTotals$`% of PF` * Positions$`% of PF`) + 
                                   (playerSeasonTotals$`% of AST` * Positions$`% of AST`) +
                                   (playerSeasonTotals$`% of BLK` * Positions$`% of BLK`),
                                 .before = '% of Min')

playerSeasonTotals <- add_column(playerSeasonTotals, "Min Adj 1" = (playerSeasonTotals$`Est Pos 1` * playerSeasonTotals$MIN + playerSeasonTotals$`Pos Num` * Positions$`Min Wt`)/(playerSeasonTotals$MIN+Positions$`Min Wt`), .before = '% of Min')

# For loop to get Trim
trim <- vector()
for (num in 1:length(playerSeasonTotals$Rank)) {
  Trim <- max(min(playerSeasonTotals[num,]$`Min Adj 1`,5),1)
 trim <- append(trim, Trim)
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Trim 1" = trim, .before = '% of Min')
#playerSeasonTotals <- add_column(playerSeasonTotals, "Trim 1" = max(min(playerSeasonTotals$`Min Adj 1`,5),1), .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Tm Avg 1" = sum(playerSeasonTotals$`Trim 1` * playerSeasonTotals$MIN, na.rm = TRUE)/TotalMinutes, .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Adj Pos 2" = playerSeasonTotals$`Min Adj 1` - (playerSeasonTotals$`Tm Avg 1` - 3), .before = '% of Min')

# For loop to get Trim
trim <- vector()
for (num in 1:length(playerSeasonTotals$Rank)) {
  Trim <- max(min(playerSeasonTotals[num,]$`Adj Pos 2`,5),1)
  trim <- append(trim, Trim)
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Trim 2" = trim, .before = '% of Min')
#playerSeasonTotals <- add_column(playerSeasonTotals, "Trim 2" = max(min(playerSeasonTotals$`Adj Pos 2`,5),1), .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Tm Avg 2" = sum(playerSeasonTotals$`Trim 2` * playerSeasonTotals$MIN, na.rm = TRUE)/TotalMinutes, .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Adj Pos 3" = playerSeasonTotals$`Min Adj 1` - (playerSeasonTotals$`Tm Avg 1` - 3) - (playerSeasonTotals$`Tm Avg 2` - 3), .before = '% of Min')

# For loop to get Trim
trim <- vector()
for (num in 1:length(playerSeasonTotals$Rank)) {
  Trim <- max(min(playerSeasonTotals[num,]$`Adj Pos 3`,5),1)
  trim <- append(trim, Trim)
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Trim 3" = trim, .before = '% of Min')
#playerSeasonTotals <- add_column(playerSeasonTotals, "Trim 3" = max(min(playerSeasonTotals$`Adj Pos 3`,5),1), .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Tm Avg 3" = sum(playerSeasonTotals$`Trim 3` * playerSeasonTotals$MIN, na.rm = TRUE)/TotalMinutes, .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Adj Pos 4" = playerSeasonTotals$`Min Adj 1` - (playerSeasonTotals$`Tm Avg 1` - 3) - (playerSeasonTotals$`Tm Avg 2` - 3) - (playerSeasonTotals$`Tm Avg 3` - 3), .before = '% of Min')

# For loop to get Position
trim <- vector()
for (num in 1:length(playerSeasonTotals$Rank)) {
  Trim <- max(min(playerSeasonTotals[num,]$`Adj Pos 4`,5),1)
  trim <- append(trim, Trim)
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Position" = trim, .before = '% of Min')
#playerSeasonTotals <- add_column(playerSeasonTotals, "Position" = max(min(playerSeasonTotals$`Adj Pos 4`,5),1), .before = '% of Min')

#Offensive Role

playerSeasonTotals <- add_column(playerSeasonTotals, "Est Off. Role 1" =
                                   OffensiveRole$Intercept + OffensiveRole$`% of ast`*playerSeasonTotals$`% of AST` + 
                                   OffensiveRole$`% of Threshold` * playerSeasonTotals$`% of ThreshPts`, .before = '% of Min')  
playerSeasonTotals <- add_column(playerSeasonTotals, "Min Adj 1_o" =
                                   (playerSeasonTotals$`Est Off. Role` * playerSeasonTotals$MIN + DefaultPos*Positions$`Min Wt`)/(playerSeasonTotals$MIN + Positions$`Min Wt`), .before = '% of Min')
# For loop to get Trim
trim <- vector()
for (num in 1:length(playerSeasonTotals$Rank)) {
  Trim <- max(min(playerSeasonTotals[num,]$`Min Adj 1_o`,5),1)
  trim <- append(trim, Trim)
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Trim 1_o" = trim, .before = '% of Min')

#playerSeasonTotals <- add_column(playerSeasonTotals, "Trim 1_o" = max(min(playerSeasonTotals$`Min Adj 1_o`,5),1), .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Tm Avg 1_o" = sum(playerSeasonTotals$`Trim 1_o` * playerSeasonTotals$MIN, na.rm = TRUE)/TotalMinutes, .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Adj Off. Role 2" = playerSeasonTotals$`Min Adj 1_o` - (playerSeasonTotals$`Tm Avg 1_o` - 3), .before = '% of Min')
# For loop to get Trim
trim <- vector()
for (num in 1:length(playerSeasonTotals$Rank)) {
  Trim <- max(min(playerSeasonTotals[num,]$`Adj Off. Role 2`,5),1)
  trim <- append(trim, Trim)
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Trim 2_o" = trim, .before = '% of Min')
#playerSeasonTotals <- add_column(playerSeasonTotals, "Trim 2_o" = max(min(playerSeasonTotals$`Adj Off. Role 2`,5),1), .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Tm Avg 2_o" = sum(playerSeasonTotals$`Trim 2_o` * playerSeasonTotals$MIN, na.rm = TRUE)/TotalMinutes, .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Adj Off. Role 3" = playerSeasonTotals$`Min Adj 1_o` - (playerSeasonTotals$`Tm Avg 1_o` - 3) - (playerSeasonTotals$`Tm Avg 2_o` - 3), .before = '% of Min')

# For loop to get Trim
trim <- vector()
for (num in 1:length(playerSeasonTotals$Rank)) {
  Trim <- max(min(playerSeasonTotals[num,]$`Adj Off. Role 3`,5),1)
  trim <- append(trim, Trim)
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Trim 3_o" = trim, .before = '% of Min')
#playerSeasonTotals <- add_column(playerSeasonTotals, "Trim 3_o" = max(min(playerSeasonTotals$`Adj Off. Role 3`,5),1), .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Tm Avg 3_o" = sum(playerSeasonTotals$`Trim 3_o` * playerSeasonTotals$MIN, na.rm = TRUE)/TotalMinutes, .before = '% of Min')
playerSeasonTotals <- add_column(playerSeasonTotals, "Adj Off. Role 4" = playerSeasonTotals$`Min Adj 1_o` - (playerSeasonTotals$`Tm Avg 1_o` - 3) - (playerSeasonTotals$`Tm Avg 2_o` - 3) - (playerSeasonTotals$`Tm Avg 3_o` - 3), .before = '% of Min')
# For loop to get Offensive Role
trim <- vector()
for (num in 1:length(playerSeasonTotals$Rank)) {
  Trim <- max(min(playerSeasonTotals[num,]$`Adj Off. Role 4`,5),1)
  trim <- append(trim, Trim)
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Offensive Role" = trim, .before = '% of Min')
#playerSeasonTotals <- add_column(playerSeasonTotals, "Offensive Role" = max(min(playerSeasonTotals$`Adj Off. Role 4`,5),1), .before = '% of Min')

# BPM Coefficients to use, based on positions

playerSeasonTotals <- add_column(playerSeasonTotals, "Adj. Pt" = (5 - playerSeasonTotals$Position)/4*BoxPlusMinusCoefficients$`Adj. Pt`[1] + (playerSeasonTotals$Position -1) / 4*BoxPlusMinusCoefficients$`Adj. Pt`[2], .before = '% of Min') 
playerSeasonTotals <- add_column(playerSeasonTotals, "FGA_bpm" = (5 - playerSeasonTotals$`Offensive Role`)/4*BoxPlusMinusCoefficients$FGA[1] + (playerSeasonTotals$`Offensive Role` -1) / 4*BoxPlusMinusCoefficients$FGA[2], .before = '% of Min') 
playerSeasonTotals <- add_column(playerSeasonTotals, "FTA_bpm" = (5 - playerSeasonTotals$`Offensive Role`)/4*BoxPlusMinusCoefficients$FTA[1] + (playerSeasonTotals$`Offensive Role` -1) / 4*BoxPlusMinusCoefficients$FTA[2], .before = '% of Min') 
#playerSeasonTotals <- add_column(playerSeasonTotals, "3PM_bpm" = (5 - playerSeasonTotals$Position)/4*BoxPlusMinusCoefficients$`3 Pt FG (bonus)`[1] + (playerSeasonTotals$Position -1) / 4*BoxPlusMinusCoefficients$`3 Pt FG (bonus)`[2], .before = '% of Min') 

it <- 4
colname <- vector()
for (coef in BoxPlusMinusCoefficients[4:12]) {
colname <- append(colname,paste0(colnames(BoxPlusMinusCoefficients[it]), "_bpm"))
playerSeasonTotals <- add_column(playerSeasonTotals, colname = (5 - playerSeasonTotals$Position)/4*coef[1] + (playerSeasonTotals$Position -1) / 4*coef[2], .before = '% of Min') 
it <- it+1
}
#Rename the columns 
colnames(playerSeasonTotals)[c(79:87)] <- colname

# Raw BPM Calculation
PositionConstantBPM <- c((-0.409*2),0,0,1.387)
PositionConstantOffensive <- c((-0.849*2),0,0,0.43)

playerSeasonTotals <- add_column(playerSeasonTotals, "Scoring" =
                                   (playerSeasonTotals$`Adj Pt` * playerSeasonTotals$`Adj. Pt` )+ 
                                   (playerSeasonTotals$FGA_p * playerSeasonTotals$FGA_bpm ) +
                                   (playerSeasonTotals$FTA_p * playerSeasonTotals$FTA_bpm )+ 
                                   (playerSeasonTotals$`3PM_p` * playerSeasonTotals$`3 Pt FG (bonus)_bpm` ),
                                 .before = '% of Min')

playerSeasonTotals <- add_column(playerSeasonTotals, "Ballhandling" =
                                   (playerSeasonTotals$AST_p * playerSeasonTotals$AST_bpm )+ 
                                   (playerSeasonTotals$TOV_p * playerSeasonTotals$TO_bpm ), .before = '% of Min')

playerSeasonTotals <- add_column(playerSeasonTotals, "Rebounding" =
                                   (playerSeasonTotals$ORB_p * playerSeasonTotals$ORB_bpm )+ 
                                   (playerSeasonTotals$DRB_p * playerSeasonTotals$DRB_bpm ) +
                                   (playerSeasonTotals$TRB_p * playerSeasonTotals$TRB_bpm ), .before = '% of Min')

playerSeasonTotals <- add_column(playerSeasonTotals, "Defense" =
                                   (playerSeasonTotals$STL_p * playerSeasonTotals$STL_bpm )+ 
                                   (playerSeasonTotals$BLK_p * playerSeasonTotals$BLK_bpm ) +
                                   (playerSeasonTotals$PF_p * playerSeasonTotals$PF_bpm ), .before = '% of Min')


pConst <- vector()
for (num in 1:length(playerSeasonTotals$Rank)) {
  if (playerSeasonTotals[num,]$Position < 3) {
    PosConst <- (playerSeasonTotals[num,]$Position - 1) / 2*PositionConstantBPM[2]+(3-playerSeasonTotals[num,]$Position )/2*PositionConstantBPM[1]
  } else {
    PosConst <- (playerSeasonTotals[num,]$Position - 3)/2*PositionConstantBPM[3]+(5-playerSeasonTotals[num,]$Position)/2*PositionConstantBPM[2]
  }
  PosConst <- PosConst + PositionConstantBPM[4] * (playerSeasonTotals[num,]$`Offensive Role`-3)
  pConst <- append(pConst,PosConst)
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Pos Const" = pConst, .before = '% of Min')

RawBPM <- vector()
for (num in 1:length(playerSeasonTotals$Rank)) {
  RawBPM <- append(RawBPM, (playerSeasonTotals[num,]$Scoring + 
                              playerSeasonTotals[num,]$Ballhandling 
                            + playerSeasonTotals[num,]$Rebounding
                            + playerSeasonTotals[num,]$Defense
                            + playerSeasonTotals[num,]$`Pos Const`))
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Raw BPM" = RawBPM, .before = '% of Min')

                                
playerSeasonTotals <- add_column(playerSeasonTotals, "Contributions" = playerSeasonTotals$`Raw BPM`*playerSeasonTotals$`% of Min`, .before = '% of Min') 

Sum <- sum(playerSeasonTotals$Contributions, na.rm = TRUE)
Tm_Adj <- (AdjTeamRating-Sum)/5


#OBPM Coefficients to use, based on Positions:
OffensiveBoxPlusMinusCoefficients
playerSeasonTotals <- add_column(playerSeasonTotals, "Adj. Pt_obpm" = (5 - playerSeasonTotals$Position)/4*OffensiveBoxPlusMinusCoefficients$`Adj. Pt`[1] + (playerSeasonTotals$Position -1) / 4*OffensiveBoxPlusMinusCoefficients$`Adj. Pt`[2], .before = '% of Min') 
playerSeasonTotals <- add_column(playerSeasonTotals, "FGA_obpm" = (5 - playerSeasonTotals$`Offensive Role`)/4*OffensiveBoxPlusMinusCoefficients$FGA[1] + (playerSeasonTotals$`Offensive Role` -1) / 4*OffensiveBoxPlusMinusCoefficients$FGA[2], .before = '% of Min') 
playerSeasonTotals <- add_column(playerSeasonTotals, "FTA_obpm" = (5 - playerSeasonTotals$`Offensive Role`)/4*OffensiveBoxPlusMinusCoefficients$FTA[1] + (playerSeasonTotals$`Offensive Role` -1) / 4*OffensiveBoxPlusMinusCoefficients$FTA[2], .before = '% of Min') 


it <- 4
colname <- vector()
for (coef in OffensiveBoxPlusMinusCoefficients[4:12]) {
  colname <- append(colname,paste0(colnames(OffensiveBoxPlusMinusCoefficients[it]), "_obpm"))
  playerSeasonTotals <- add_column(playerSeasonTotals, colname = (5 - playerSeasonTotals$Position)/4*coef[1] + (playerSeasonTotals$Position -1) / 4*coef[2], .before = '% of Min') 
  it <- it+1
}


#Rename the columns 
colnames(playerSeasonTotals)[c(98:106)] <- colname

PositionConstantOffensive <- c((-0.849*2),0,0,0.43)

playerSeasonTotals <- add_column(playerSeasonTotals, "Scoring_o" =
                                   (playerSeasonTotals$`Adj Pt` * playerSeasonTotals$`Adj. Pt_obpm` )+ 
                                   (playerSeasonTotals$FGA_p * playerSeasonTotals$FGA_obpm ) +
                                   (playerSeasonTotals$FTA_p * playerSeasonTotals$FTA_obpm )+ 
                                   (playerSeasonTotals$`3PM_p` * playerSeasonTotals$`3 Pt FG (bonus)_obpm` ),
                                 .before = '% of Min')

playerSeasonTotals <- add_column(playerSeasonTotals, "Ballhandling_o" =
                                   (playerSeasonTotals$AST_p * playerSeasonTotals$AST_obpm )+ 
                                   (playerSeasonTotals$TOV_p * playerSeasonTotals$TO_obpm ), .before = '% of Min')

playerSeasonTotals <- add_column(playerSeasonTotals, "Rebounding_o" =
                                   (playerSeasonTotals$ORB_p * playerSeasonTotals$ORB_obpm )+ 
                                   (playerSeasonTotals$DRB_p * playerSeasonTotals$DRB_obpm ) +
                                   (playerSeasonTotals$TRB_p * playerSeasonTotals$TRB_obpm ), .before = '% of Min')

playerSeasonTotals <- add_column(playerSeasonTotals, "Defense_o" =
                                   (playerSeasonTotals$STL_p * playerSeasonTotals$STL_obpm )+ 
                                   (playerSeasonTotals$BLK_p * playerSeasonTotals$BLK_obpm ) +
                                   (playerSeasonTotals$PF_p * playerSeasonTotals$PF_obpm ), .before = '% of Min')


pConst <- vector()
for (num in 1:length(playerSeasonTotals$Rank)) {
  if (playerSeasonTotals[num,]$Position < 3) {
    PosConst <- (playerSeasonTotals[num,]$Position - 1) / 2*PositionConstantOffensive[2]+(3-playerSeasonTotals[num,]$Position )/2*PositionConstantOffensive[1]
  } else {
    PosConst <- (playerSeasonTotals[num,]$Position - 3)/2*PositionConstantOffensive[3]+(5-playerSeasonTotals[num,]$Position)/2*PositionConstantOffensive[2]
  }
  PosConst <- PosConst + PositionConstantOffensive[4] * (playerSeasonTotals[num,]$`Offensive Role`-3)
  pConst <- append(pConst,PosConst)
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Pos Const_obpm" = pConst, .before = '% of Min')

RawOBPM <- vector()
for (num in 1:length(playerSeasonTotals$Rank)) {
  RawOBPM <- append(RawOBPM, (playerSeasonTotals[num,]$Scoring_o + 
                              playerSeasonTotals[num,]$Ballhandling_o 
                            + playerSeasonTotals[num,]$Rebounding_o
                            + playerSeasonTotals[num,]$Defense_o
                            + playerSeasonTotals[num,]$`Pos Const_obpm`))
}
playerSeasonTotals <- add_column(playerSeasonTotals, "Raw OBPM" = RawOBPM, .before = '% of Min')


playerSeasonTotals <- add_column(playerSeasonTotals, "Contrib" = playerSeasonTotals$`Raw OBPM`*playerSeasonTotals$`% of Min`, .before = '% of Min') 



Sum_o <- sum(playerSeasonTotals$Contrib, na.rm = TRUE)
Tm_Adj_o <- (AdjOffRating-Sum_o)/5
totals <- rbind(totals, playerSeasonTotals)

#Finish Calculations on Separate Dataframe

BPM <- data.frame(
             playerSeasonTotals$Name, 
             playerSeasonTotals$Team,
             playerSeasonTotals$Position, 
             playerSeasonTotals$`Offensive Role`, 
             playerSeasonTotals$MIN, 
             playerSeasonTotals$MIN/playerSeasonTotals$GP,
             playerSeasonTotals$`Raw BPM`+Tm_Adj,
             playerSeasonTotals$`Raw OBPM`+Tm_Adj_o)

BPM <- add_column(BPM, "DBPM" = BPM$playerSeasonTotals..Raw.BPM....Tm_Adj - BPM$playerSeasonTotals..Raw.OBPM....Tm_Adj_o)
BPM <- add_column(BPM, "Contribution" = BPM$playerSeasonTotals..Raw.BPM....Tm_Adj*playerSeasonTotals$`% of Min`)
BPM <- add_column(BPM, "VORP" = (BPM$playerSeasonTotals..Raw.BPM....Tm_Adj + 2)*playerSeasonTotals$`% of Min`*TeamGames/20)
BPM <- add_column(BPM, "eFG" = (playerSeasonTotals$FGM+(0.5*playerSeasonTotals$`3PM`))*100/playerSeasonTotals$FGA)
BPM <- add_column(BPM, "TOV" = ((playerSeasonTotals$TOV*100)/(playerSeasonTotals$TOV+(0.44*playerSeasonTotals$FTA)+playerSeasonTotals$FGA)))
BPM <- add_column(BPM, "FTF" = (playerSeasonTotals$FTM*100/playerSeasonTotals$FGA))
BPM <- add_column(BPM, "PPP" = round(playerSeasonTotals$PTS/playerSeasonTotals$Possessions*5,3))

colnames(BPM)[c(1:9)] <- c("Name", "Team", "Position", "Offensive_Role", "Minutes", "MPG", "BPM", "OBPM", "DBPM")

BPM <- cbind(BPM$Name, BPM$Team, round(BPM[c(3:14)],1), BPM$PPP)

colnames(BPM)[c(1,2,15)] <- c("Name", "Team", "PPP")


league <- rbind(league, BPM)
}


## Complete PER Calculation, uses possession estimator instead of actual possessions

qualifiedPlayers <- league
qualifiedTotals <- totals

lgPace <- (lgPoss * 5) / lgMin

lgVOP <- lgPTS / (lgFGA - lgORB + lgTO + (0.44*lgFTA))
lgDRBP <- (lgTRB - lgORB) / lgTRB
lgFactor <- ((2/3) - ((0.5* (lgAST/lgFG))/(2*(lgFG/lgFT))))

lguPER <- (1/lgMin)*(lgThree+((2.0 * lgAST)/3.0)+((2-(lgFactor*(lgAST/lgFG)))*lgFG) + (0.5*lgFT*(2-(1.0/3.0)*(lgAST/lgFG))) - (lgVOP*lgTO) - (lgVOP*lgDRBP*(lgFGA - lgFG)) - (lgVOP*0.44*(0.44+(0.56*lgDRBP))*(lgFTA-lgFT))+(lgVOP*(1-lgDRBP)*(lgTRB-lgORB))+(lgVOP*lgDRBP*lgORB)+(lgVOP*lgSTL)+(lgVOP*lgDRBP*lgBLK)-(lgPF*((lgFT/lgPF)-0.44*(lgFTA/lgPF)*lgVOP)))


PERlist <- vector()
colnames(TeamStats) <- c("Team", "GP", "MIN", "PTS", "FG", "FGA", "FG%", "X3FG", "X3FGA", "X3FG%", "FT", "FTA", "FT%", "OFF", "DEF", "TOT", "A", "STL", "BLK", "TO", "PF", "POSS", "oppPOSS", "Pace")
TeamStats <- TeamStats %>% mutate_at(c(2:24), as.numeric)
for (l in 1:length(qualifiedPlayers$Name)) {

  uPER <-  (1/qualifiedTotals$MIN[l])*(qualifiedTotals$`3PM`[l]+((2.0 * qualifiedTotals$AST[l])/3.0)
    +((2-(lgFactor*((TeamStats$A[which(qualifiedPlayers$Team[l] == TeamStats$Team)])/(TeamStats$FG[which(qualifiedPlayers$Team[l] == TeamStats$Team)]))))*qualifiedTotals$FGM[l]) + (0.5*qualifiedTotals$FTM[l]*(2-(1.0/3.0)*((TeamStats$A[which(qualifiedPlayers$Team[l] == TeamStats$Team)])/(TeamStats$FG[which(qualifiedPlayers$Team[l] == TeamStats$Team)]))))
    - (lgVOP*qualifiedTotals$TOV[l]) - (lgVOP*lgDRBP*(qualifiedTotals$FGA[l] - qualifiedTotals$FGM[l])) 
    - (lgVOP*0.44*(0.44+(0.56*lgDRBP))*(qualifiedTotals$FTA[l]-qualifiedTotals$FTM[l]))
    +(lgVOP*(1-lgDRBP)*(qualifiedTotals$TRB[l]-qualifiedTotals$OFF[l]))
    +(lgVOP*lgDRBP*qualifiedTotals$OFF[l])+(lgVOP*qualifiedTotals$STL[l])
    +(lgVOP*lgDRBP*qualifiedTotals$BLK[l])
    -(qualifiedTotals$PF[l]*((lgFT/lgPF)-0.44*(lgFTA/lgPF)*lgVOP)))
  
  PER <- round((uPER * (lgPace/TeamStats$Pace[which(qualifiedPlayers$Team[l] == TeamStats$Team)]))*(15/lguPER),1)
  
  PERlist <- append(PERlist, PER)
  
}
qualifiedPlayers <- add_column(qualifiedPlayers, "PER" = PERlist)

qualifiedPlayers <- qualifiedPlayers[order(qualifiedPlayers$VORP, decreasing = TRUE),]
rownames(qualifiedPlayers) <- NULL
kable(qualifiedPlayers[1:10,c(1,2,7,11,12,13,14,15,16)], caption = paste0("Advanced Player Metrics for the Top 10 CEBL Players During the ", yr, " Season, per VORP"))
options(warn = oldw)
```


# 3.2 Application to the CEBL
Table 3.1 displays the advanced statistics of the top ten CEBL players, as per VORP. Due
to the relative infancy of the CEBL, these calculations are the first publicly available set of
their kind [1]. There are many interesting discussions that arise from viewing this table, and
one example is the power of VORP over BPM and PER in terms of strict player valuation.
Take Hamilton’s Christian Vital, who led the CEBL in BPM by a considerable margin and
was second in both PER and PPP but was credited as the fourth on the VORP total and
wasn’t even shortlisted for the league’s most valuable player honors. This is because Vital
only appeared in 13 out his team’s 20 regular season games, and this through no fault of his
own - part of his time away was spent playing for the Toronto Raptors during the 2022 NBA
Summer League. BPM and PER, which are rate statistics, correctly identify Vital as a prolific
producer when on the court, and it comes as no surprise that upon returning he led his team
to the CEBL championship and was named finals MVP. However, since his overall impact over
the full course of the season was not felt to his team as much as the players ranked ahead of
him (Carr, 18 games played, Gibson 19, Ahmad 20), VORP accurately definition his value in
terms of overall contribution to his team’s success.

When comparing the players from Table 3.1 to their league award considerations, a couple
of significant omissions can be found that demonstrate the incompleteness of valuations based
on traditional counting statistics. Montreal’s Isiah Osborne, who was left off of both Sixth
Man of the Year and Top Canadian ballots, was second only to the aforementioned Christian
Vital in BPM and led the CEBL with a TOV% of 4.5%, nearly half that of the next closest
player (E.J. Onu, 8.5%). This mistake-free and efficient style of basketball is highly valued by
statistical metrics but unfortunately caused Osborne to be overlooked by more prolific point
accumulators. An above average interior scorer, Osborne averaged a “pedestrian ”10.3 points
per game while playing for the league’s worst team, and these factors led to him not being
properly credited for the impact he carried. Similarly, Edmonton’s Jordan Baker was left off of
the Top Canadian ballot even though he led all Canadian-born players in VORP and was top-5
among Canadians in both BPM and PER. Baker is the CEBL’s all-time leader in rebounds
and assists and had an impressive 2022 campaign, acting as a major focal point of his team’s
offensive schemes even though he often wasn’t the main scoring option and recorded a PPP
total reflective of that. His undervaluation comes partially due to his unseen defensive impact
and high creation output, which, though not seen by basic counting statistics and some of the
offensive top-down statistics, were recognized accordingly by advanced statistics.

# 4 Conclusion
Advanced statistical methods utilizing regressions and Bayesian-based constructs have been
used for nearly 30 years in the NBA to enhance the quantitative understanding of the game of
basketball. This paper investigated possession-based metrics and their applications within U
SPORTS and the CEBL to assess their efficacy when addressing questions of how to evaluate
team and individual performance. It has been demonstrated that these analytical instruments
maintain excellent functionality when applied to data from such leagues, with models utilizing
the four factors alone explaining up to 98% of a team’s variation in win percentage for U
SPORTS men’s and women’s teams. Expected win percentage predictions perform extremely
well, with low error rates and R^2^ values near or above 87% for the 2021-2022 U SPORTS
regular season. Established individual player metrics such as Box Plus Minus, Value Over
Replacement Player, and Player Efficiency Rating were applied to CEBL players to illustrate
oversights in traditional player valuations and provide observers with a better ranking system.
The introduction of a publicly accessible repository and website to access this data allows both
fans and decision makers to utilize them to maximize and understand their team’s success.
Advances in these areas and the inclusion of more descriptive data types such as spatial and
play-by-play will ensure that Canadian Basketball leagues continue to produce high-quality
basketball as it grows.

# 4.1 Future Works

There are limitations to the individual player evaluation applications from Section 3 due to lack
of information or available data surrounding their original regression and plus-minus methods.
In PER’s case, a low-level description of the coefficients that compute the league value of an
offensive possession and the importance of each player’s contribution is not publicly available.
Therefore, while it is effective in general for intra-league player comparisons, its values become
increasingly approximate the farther away the style of play of the league is from that of the
NBA. Box Plus Minus tolerates different league contexts much better by comparing players to
their own teammates before adjusting based on individual statistics but struggles to properly
adjust coefficients in leagues that have widely different PPP or possession-ending event type
distributions. For example, differing offensive rebounding distributions and large eFG% discrepancies
that are apparent between U SPORTS men and women should result in a different
coefficient weights given to each contribution. Therefore, attaining the additional information
and data necessary to recreate BPM, PER, and similar catch-all metrics from the ground up
will improve our notion of player value within U SPORTS and the CEBL.

# 4.2 Supplementary Material
The data and code necessary to reproduce the analyses performed in this paper are available
from GitHub at https://github.com/awosoga/CMSAC_2022.

# References
[1] David Awosoga. Canadian basketball analytics. https://canadianbasketballanalytics.carrd.co. Accessed: 2022-08-12.

[2] Jeremias Engelmann. Possession-based player performance analysis in basketball (adjusted
+/- and related concepts). 2016.

[3] Justin Kubatko, Dean Oliver, Kevin Pelton, and Dan Rosenbaum. A starting point for
analyzing basketball statistics. Journal of Quantitative Analysis in Sports, 3:1–1, 02 2007.

[4] Austin Clemens. Nylon calculus 101: Plus-minus and adjusted plus-minus. https://fansided.com/2014/09/25/glossary-plus-minus-adjusted-plus-minus/. Accessed:
2022-02-26.

[5] Daniel Myers. About box plus/minus (bpm). https://www.basketball-reference.com/about/bpm2.html. Accessed: 2022-02-21.

[6] Krishna Narsu and Tim/Cranjis McBasketball. Lebron introduction. https://www.bball-index.com/lebron-introduction/. Accessed: 2022-03-01.

[7] Nate Silver. Introducting raptor, our new metric for
the modern nba. https://fivethirtyeight.com/features/introducing-raptor-our-new-metric-for-the-modern-nba/. Accessed: 2022-03-01.

[8] Petar Vračar, Erik Štrumbelj, and Igor Kononenko. Modeling basketball play-by-play
data. Expert Systems with Applications, 44:58–66, 2016.

[9] Zachary Terner and Alexander Franks. Modeling player and team performance in basketball,
2020.

[10] Alexander Franks, Andrew Miller, Luke Bornn, and Kirk Goldsberry. Characterizing
the spatial structure of defensive skill in professional basketball. The Annals of Applied
Statistics, 9, 05 2014.

[11] Edgar Santos Fernandez, Paul Wu, and Kerrie Mengersen. Bayesian statistics meets
sports: a comprehensive review. Journal of Quantitative Analysis in Sports, 15(4):289–
312, 2019.

[12] U SPORTS. U sports. https://usports.ca/en. Accessed: 2022-02-21.

[13] PrestoSports. Prestosports | all-in-one sports technology platform. https://www.prestosports.com/landing/index. Accessed: 2022-02-21.

[14] Martin Timmerman. U sports hoops, university baskbetball in canada. https://usportshoops.ca. Accessed: 2022-02-21.

[15] RealGM. Canadian elite basketball league. https://basketball.realgm.com/international/league/128/Canadian-Elite-Basketball-League/. Accessed: 2022-
08-01.

[16] R Core Team. R: A Language and Environment for Statistical Computing. R Foundation
for Statistical Computing, Vienna, Austria, 2022.

[17] Dean Oliver. Basketball on paper : rules and tools for performance analysis / Dean Oliver.
Brassey’s, Inc., Washington, D.C, 1st ed. edition, 2004.

[18] T.M. Tango, M.G. Lichtman, and A.E. Dolphin. The Book: Playing the Percentages in
Baseball. Potomac Books, 2007.

[19] R. Carter Hill, William E Griffiths, and G. C. (Guay C.) Lim. Principles of econometrics
/ R. Carter Hill, William E. Griffiths, Guay C. Lim. Wiley, Hoboken, fifth edition, 2018.

[20] Marek Hlavac. stargazer: Well-Formatted Regression and Summary Statistics Tables. R
package version 5.2.3, 2022.

[21] Tarek Al Baghal. Are the four factors indicators of one factor? an application of structural
equation modeling methodology to nba data in prediction of winning percentage. Journal
of Quantitative Analysis in Sports, 8, 01 2012.

[22] Konstantinos Kotzias. The four factors of basketball as a measure of success. https://statathlon.com/four-factors-basketball-success/. Accessed: 2022-05-09.

[23] Duane Rockerbie. The Economics of Professional Sports 2017. 11 2017.

[24] Synergy Sports Technology. Synergy sports. https://synergysports.com. Accessed:
2022-04-18.

[25] J. Hollinger. Pro Basketball Forecast. PRO BASKETBALL PROSPECTUS. Potomac
Books, Incorporated, 2005.

[26] Basketball Reference. Calculating per. https://www.basketball-reference.com/about/per.html. Accessed: 2022-08-12.

[27] Daniel Myers. Box plus/minus 2.0 example calculations. https://docs.google.com/spreadsheets/d/1PhD9eo3IqzpQo21-yVJPQzYjpXl_h-ZonIKqGEKBqwY/edit#gid=307166562. 
Accessed: 2022-08-15.

[28] Bryan Kalbrosky. What is the best advanced statistic for basketball?
nba executives weigh in. https://hoopshype.com/lists/advanced-stats-nba-real-plus-minus-rapm-win-shares-analytics/. Accessed:
2022-03-01.
